A Large-scale Study of Polymorphism in R

### 2019-03-22 paper/main.tex
put in plots, i will tinker with them over the next bit
counts in top 10 table, some stuff left that i will get once data avail

@@

In this section, we present our dataset. R is designed to accommodate data
analysts.  Typical workflows go from data import, to data cleaning, followed
by steps of modeling, transformation and visualization. Often the code that
implement these analysis pipelines together with their results resides in
notebooks. Unfortunately notebooks are typically not shared in public
repositories such as GitHub, and if they were, the data required to run them
would surely not be shared. For this reason our analysis focuses on
packages; these bundle reusable units of R code with documentation, sample
data, use-cases and tests.

We focus on packages hosted on the \emph{Comprehensive R Archive Network} or
CRAN.  With over 13,000 packages, CRAN is the largest repository of software
written in R. It is experiencing sustained growth with an average of size
new packages a day~\cite{LIgges2017}.  Unlike sites like GitHub, CRAN is a
curated collection of packages. A package is only accepted to CRAN if it
abides by a number of well-formedness rules.  Most relevant for our
purposes, packages must have data, examples, vignettes and tests. All of
which must successfully run. From our perspective this means that each of
the package in CRAN comes with up to a dozen executable scripts that
exercise some of its functionality.

Our corpus is a subset of CRAN, consisting of \PACKAGES packages, accounting
for some \PERCENTCRAN of all packages.  These packages have a total of \RLOC
lines of R code and \CLOC lines of C code. For each package, we extracted
all executable code snippets from documentation, vignettes and tests and ran
them independently recording all call to R functions.  It is noteworthy that
in order to run the scripts in one package, it is often necessary to load a
number of other packages.  In~\cite{issta18}, the authors estimated code
coverage to be around 68\% when including reverse dependencies.  The
rationale for our choice of packages was pragmatic, our infrastructure adds
overhead to the execution of scripts, we eliminated any package that would
take more than one hour to run. We expect to extend our coverage of CRAN for
the next version of this paper.

@@

% https://www.r-pkg.org/downloaded
\begin{figure}[!th]{\footnotesize\begin{tabular}{@{}r||l|r|r|r@{}}\hline
\bf Package & \bf Description & \bf R LOC &\bf C LOC &\bf Scripts\\\hline
\tt Rcpp  & Seamless C++ integration & 10K & 3K & 12 \\
\tt rlang & Functions for 'Tidyverse'&&& \\
\tt glue  & Interpreted string literals&&& \\
\tt tibble & Simple data frames&& &\\
\tt stringi &  String processing &&& \\
\tt ggplot2 & Data visualisations&& &\\
\tt dplyr  &  Data manipulation&&& \\
\tt pillar & Formatting for columns&&& \\
\tt R6 & Classes w. ref. semantics&&& \\
\tt stringr & String operations&&& \\
\end{tabular}}\caption{10 Most Downloaded Packages.}\label{most}
\end{figure}

>>>>

\graphicspath{ {./plots/} }

@@

In this section, we present our dataset. R is designed to accommodate data
analysts.  Typical workflows go from data import, to data cleaning, followed
by steps of modeling, transformation and visualization. Often, the code that
implements these analysis pipelines together with their results resides in
notebooks. Unfortunately notebooks are typically not shared in public
repositories such as GitHub, and if they were, the data required to run them
would surely not be shared. For this reason our analysis focuses on
packages; these bundle reusable units of R code with documentation, sample
data, use-cases and tests.

We focus on packages hosted on the \emph{Comprehensive R Archive Network} or
CRAN.  With over 13,000 packages, CRAN is the largest repository of software
written in R. It is experiencing sustained growth with an average of size
new packages a day~\cite{LIgges2017}.  Unlike sites like GitHub, CRAN is a
curated collection of packages: A package is only accepted to CRAN if it
abides by a number of well-formedness rules.  Most relevant for our
purposes, packages must have data, examples, vignettes and tests, all of
which must successfully run. From our perspective this means that each package 
in CRAN comes with several executable scripts that exercise some of its functionality.
Notable exceptions to this rule are packages only containing data, which have no runnable
code but are referenced by other packages \AT{I should collect stats on this}.

Our corpus is a subset of CRAN, consisting of \PACKAGES packages, accounting
for some \PERCENTCRAN of all packages.  These packages have a total of \RLOC
lines of R code and \CLOC lines of C code. For each package, we extracted
all executable code snippets from documentation, vignettes and tests and ran
them independently recording all calls to R functions.  It is noteworthy that
in order to run the scripts in one package, it is often necessary to load a
number of other packages.  In~\cite{issta18}, the authors estimated code
coverage to be around 68\% when including reverse dependencies.  As our infrastructure
adds overhead to script execution, coupled with the fact that some scripts take 
inordinate amounts of time to run, we limit our analysis of any given package to one hour.
We capture and record function signatures on-the-fly, though, so we do collect (partial) data even 
for these long-running packages. \AT{We expect to extend our coverage of CRAN for the next version of this paper.}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.6]{linesofrandccode.pdf}
\caption{Lines of Code (R and C) in the Analyzed Corpus}
\label{allcloc}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.6]{recordsbypkg}
\caption{Numbers of Recorded Function Invocations in the Analyzed Corpus}
\label{allcloc}
\end{center}
\end{figure}

% https://www.r-pkg.org/downloaded
\begin{figure}[!th]{\footnotesize\begin{tabular}{@{}r||l|r|r|r@{}}\hline
\bf Package & \bf Description & \bf R LOC &\bf C LOC &\bf Scripts\\\hline
\tt Rcpp  & Seamless C++ integration & 2.1K & 4K & 12 \\
\tt rlang & Functions for 'Tidyverse' & 4.4K & 3.8K & ? \\
\tt glue  & Interpreted string literals & 0.2K & 0.2K & 8 \\
\tt tibble & Simple data frames & 1.3K & 0.3K & ? \\
\tt stringi &  String processing & 1.5K & 515K & ? \\
\tt ggplot2 & Data visualisations & 10.8K & 0 & 130 \\
\tt dplyr  &  Data manipulation & 3.8K & 4.7K & 78 \\
\tt pillar & Formatting for columns & 1.3K & 0 & ? \\
\tt R6 & Classes w. ref. semantics & 0.6K & 0 & ? \\
\tt stringr & String operations & 0.5K & 0 & ? \\
\end{tabular}}\caption{10 Most Downloaded Packages.}\label{most}
\end{figure}
